Anomalous waves triggered by abrupt depth changes: experiments and truncated KdV theory

### 2019-10-09 paper2_JNLS/paper.tex
NM

	The paddle amplitude in Fig.~\ref{ExpDiagStats} is $\Dphi = 1.38^{\circ}$, and this value was varied systematically in the range $\Dphi = 0.125^{\circ}$--$2^{\circ}$ to probe the different regimes of wave behavior, from linear to strongly nonlinear waves. Figure \ref{ExpSpatialStats} shows spatial statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ for six different driving amplitudes (see legend). For each quantity, we show the standard deviation, the skewness, and the (excess) kurtosis as they vary in the vicinity of the ADC, located at $x=0$. First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves. Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC. In experiments with sufficiently high driving amplitude, though, the skewness and kurtosis (b)--(c) respond strongly to the ADC. In these experiments, both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm. Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude. The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e). Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution. We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.

@@

	Figure \ref{uhist} shows displacement histograms from these TKdV simulations in comparison to the experiment histograms. The later representative exactly the same data as shown in Fig.~\ref{ExpDiagStats}, only converted to the dimensionless variables. The TKdV simulations do a remarkable job of capturing both the symmetric upstream statistics and the skewed downstream statistics. Amazingly, the simulations reproduce the measured distributions in great detail, both in the shape of the distribution and the scale.
%
This comparison offers convincing evidence that the theory works and the parameters have been matched as well as possible. \nick{Reword}

@@

	These slope histograms show intriguing similarities and differences. First, the upstream slope distribution is captured well by the TKdV simulations, both in its nearly symmetric shape and its scale. We note that, while the standard deviation of displacement was input into the TKdV theory, the slope standard deviation was not. Downstream, both the experimental and theoretical slope distributions  spread significantly. Interestingly, in the experiments, it is not the standard deviation of slope that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but the excess kurtosis, which jumps from a value of 0.46 to 3.4. Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50. Thus, the jump in kurtosis, while not as extreme as that measured in experiments, is quite significant in the theory. These elevated levels of kurtosis give the distributions the look of being stretched in the tails and pointy near the centers (especially in the experiments).

@@

We show individual trajectories in the next figure for upstream and downstream. In the upstream flow, due to the coefficients, the nonlinearity is very weak. Hence, wave propagation is dominated by linear dispersion. Indeed, with weak nonlinearity, \eqref{dimlessKdV} is approximated by
\begin{equation}
u_t + C_2 \drat^{1/2} \, u_{xxx} = 0
\qquad \text{for } x \in [-\pi,\pi] \\
\end{equation}
which implies the dispersion relation $\omega_k = -C_2 k^3$ or for phase velocity $c_k = -C_2 k^2$, which implies left-going waves \cite{majdaqi2019}.
% Further details, see page 6 of Majda Qi, J Stat Phys.

>>>>

	The paddle amplitude in Fig.~\ref{ExpDiagStats} is $\Dphi = 1.38^{\circ}$, and this value was varied systematically in the range $\Dphi = 0.125^{\circ}$--$2^{\circ}$ to probe the different regimes of wave behavior, from linear to strongly nonlinear waves. Figure \ref{ExpSpatialStats} shows long-time statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ as they vary in space for six different driving amplitudes (see legend). We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
\begin{align}
& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
& \std(q) = q_{std}= \sqrt{\var(q)}
&&\mbox{\em standard deviation} \\
& \skw(q) = \frac{\mean{q^3}} {q_{std}^3}	
&&\mbox{\em skewness} \\
& \kurt(q) = \frac{\mean{q^4}} {q_{std}^4} - 3
&&\mbox{\em (excess) kurtosis}
\end{align}
where $\mean{}$ indicates a mean (here a long-time mean). Hereafter, we will simply refer to the excess kurtosis as kurtosis.

	Figure \ref{ExpSpatialStats} shows how these statistics vary in the vicinity of the ADC, located at $x = 0$, for both displacement and slope. First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves. Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC. In experiments with sufficiently high driving amplitude, though, the skewness and kurtosis (b)--(c) respond strongly to the ADC. In these experiments, both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm. Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude. The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e). Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution. We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.

@@

	Figure \ref{uhist} shows displacement histograms from these TKdV simulations in comparison to the experiment histograms. The later representative exactly the same data as shown in Fig.~\ref{ExpDiagStats}, only converted to the dimensionless variables. The TKdV simulations do a remarkable job of capturing both the symmetric upstream statistics and the skewed downstream statistics. Amazingly, the simulations reproduce the measured distributions in great detail, both in the shape of the distribution and the scale.
%
This comparison offers convincing evidence that the theory works and the parameters have been matched as well as possible. \nick{Reword and expand}

@@

	These slope histograms show intriguing similarities and differences. First, the upstream slope distribution is captured well by the TKdV simulations, both in its nearly symmetric shape and its scale. We note that, while the standard deviation of displacement was input into the TKdV theory, the slope standard deviation was not. Downstream, both the experimental and theoretical slope distributions  spread significantly. Interestingly, in the experiments, it is not the standard deviation of slope that increases significantly upon crossing the depth change ($\std(\eta_x)$ increases from 0.15 to 0.17), but the excess kurtosis that jumps from a value of 0.46 to 3.4. Likewise, in the TKdV theory, $\std(\eta_x)$ grows from 0.13 upstream to 0.33 downstream, and $\kurt(\eta_x)$ grows from -0.04 to 0.50. Thus, the jump in kurtosis, while not as extreme as that measured in experiments, is quite significant in the theory. These elevated levels of kurtosis give the distributions the look of being stretched in the tails and pointy near the centers (especially in the experiments).

@@

Next, to complement the statistical comparison, we investigate some individual solution trajectories from the TKdV simulations.

\vsp{5}
We show individual trajectories in the next figure for upstream and downstream. In the upstream flow, due to the coefficients, the nonlinearity is very weak. Hence, wave propagation is dominated by linear dispersion. Indeed, with weak nonlinearity, \eqref{dimlessKdV} is approximated by
\begin{equation}
u_t + C_2 \drat^{1/2} \, u_{xxx} = 0
\qquad \text{for } x \in [-\pi,\pi] \\
\end{equation}
which implies the dispersion relation $\omega_k = -C_2 k^3$ or for phase velocity $c_k = -C_2 k^2$, which implies left-going waves \cite{majdaqi2019}.
% Further details, see page 6 of Majda Qi, J Stat Phys.
